---
title: "HW2: Titis in the Mist"
author: "maekell98"
date: "9/21/2021"
output: html_document
---

Every Saturday, at the same time, a primatologist goes and sits in the forest in the morning and listens for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. Based on previous knowledge, she believes that the mean number calls she will hear in that time is exactly 15. Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.

#**Part 1**
What is the probability that she will hear more than 8 calls during any given session?
```{r probabilty >8 calls in a session}
ppois(8, 15, lower = FALSE)
```
#[1] 0.9625535 -- 96% probability of having >8 calls in a session
#x = 8, lambda = 12 (the average per time frame), and lower=FALSE means this calculates the upper tail aka probability of having 8 or more calls in the time frame


#**Part 2**
What is the probability that she will hear no calls in a session?
```{r no calls in a session}
ppois(0, 15)
```
#[1] 3.059023e-07 -- probability of zero calls in a session

#**Part 3**
What is the probability that she will hear exactly 3 calls in a session?
```{r exactly 3 calls in a session}
dpois(3, 15)
```
#[1] 0.0001720701 is probability of exactly 3 calls in a session.
#I first used ppois and just did ppois(3, 15)-ppois(2, 15) which got the correct answer, but then I realized what dpois was for haha....On that note, qpois is for calcuating a number of events occuring based on a specified cumulative probability. 


#**Part 4**
Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.
```{r Poisson mass function 0 ≤ x ≤ 30}
x <- 0:30
l = 15
probset <- dpois(x = x, lambda = l)
barplot(probset, names.arg = x, space = 0, xlab = "x", ylab = "Pr(X = x)", main = "Probability Mass Function")
```


#**Part 5**
Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).
```{r simulate 104 results from distribution}
rp <- rpois(104, 15)
rp
```
#rpois simulates random numbers from the Poisson distribution
#  [1] 11 19 13 13 12 23 11 14 12 18 19 12 15 17 14 20 14[18]  9 20  9 14  8 11 15 15 19 14 14 17 14 15 14 14 14[35] 15 20 11 18 10 11 18 11 16 10 13 14 16 12 23 16 21[52] 16 22 14 20 11 16 11 15 14 16 13 10 13 15 11 13 15[69] 12 17 19 12 16 12 21 15 10 21 18 17 15 20 20 19 14[86] 13 13 14 14 16 19 18 20 22 16 11 11 23 17 18 20 12[103] 11 18


#**Part 6**
Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?
```{r plot results as histogram}
Simulated_Results <- rpois(104, 15)
hist(x = Simulated_Results, xlim = c(0,30), col = "darkseagreen1", border = "darkseagreen", ylab = "Frequency", xlab = "Number of Titi Calls")
```
#The shape of the histogram is similar to the shape of the probability mass function, but the PMF is symmetric and the histogram is skewed right. This is because the histogram is constructed from randomly generated numbers based on the probability mass function, so it's possible that random chance can lead to asymmetric results. If you increase the number of randomly generated results, the histogram starts to resemble the PMF more. 
#xlim uses c(...,...) to specify the lower and upper horizontal limits. 
